---
title: "Descriptor Calculation"
author: "Egon"
date: "24 september 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(itertools)
library(rcdk)
```

# Descriptor Calculation with the RCDK

First we read the compounds into a table:

```{r}
assayData = read.table(
  file='/home/egonw/bioclipse-workspace/MSB1015/aid624202.csv',
  sep=',', header=TRUE, skip=5
)
```

We are only interested only in some of the columns with the PubChem Substance ID ("X") and
the PubChem Activity Score ("X.3").

```{r}
qsarData = assayData[,c("X","X.3")]
colnames(qsarData) = c("SID","Score")
```

Many compounds are inactive, i.e. have a zero score:

```{r}
inactives = which(qsarData[,"Score"] == 0)
actives = which(qsarData[,"Score"] != 0)
length(inactives)
```

But we a more balanced data set and will force this by selecting about 700 actives and about
350 inactives:

```{r}
selectedInactives = sample(inactives, 350)
selectedActives = sample(actives, 700)
qsarSmallData = rbind(qsarData[selectedActives,], qsarData[selectedInactives,])
```

Next, we will process the Structure Data File (SDF) with all the chemical structures and
calculate descriptors for the selected compounds. For this we will use the `rcdk` package
and an iterating reader, so that we do not clog memory:

```{r}
selectedSusbtanceIDs = qsarSmallData[,"SID"]
iter <- iload.molecules('/home/egonw/bioclipse-workspace/MSB1015/aid624202.sdf', type='sdf')
pb <- txtProgressBar(min = 0, max = 377550, style = 3)
molsDone = 0
hitsFound = 0
mols = list()
while(hasNext(iter)) {
  mol <- nextElem(iter)
  sdfSID = get.property(mol, "PUBCHEM_SUBSTANCE_ID")
  if (sdfSID %in% selectedSusbtanceIDs) {
    hitsFound = hitsFound + 1
    mols[[hitsFound]] = mol
  }
  molsDone = molsDone + 1
  setTxtProgressBar(pb, molsDone)
}
close(pb)
```

Now it's time to calculate descriptors for these 1050 compounds. We first define the list
of descriptors we want to calculate:

```{r}
descNames <- unique(unlist(sapply(get.desc.categories()[2:4], get.desc.names)))
```

And then calculate the descriptors for all compounds:

```{r}
descs <- eval.desc(mols, descNames, verbose = TRUE)
class(descs)
```

Let's write the descriptors to a CSV file, but we want :

```{r}
molSIDs = unlist(lapply(mols, function(x) { get.property(x, "PUBCHEM_SUBSTANCE_ID")} ))
write.csv(cbind(molSIDs, descs), file = "/home/egonw/bioclipse-workspace/MSB1015/descriptors.csv")
```

That's it for today.
